---
title: "PH1930 Final Project"
author: "Group 2: Yu Bin Chen, Rachit Sabharwal, Caroline Schaefer, Pagna Sok"
date: "2023-12-10"
output:
  html_document: default
  pdf_document: default
---

This report contains results of minimizing the equation using the coordinate descent (CD) and alternating direction method of multipliers (ADMM). Functions are run using C++ code and implemented with the R packages Rcpp and RcppArmadillo: 

\begin{align*}
\frac{1}{2}u^TAu-ub^T+\lambda ||u||_1\\
\end{align*}  


```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("mycglassoG2_0.1.0.tar.gz")

library(mycglassoG2)
library(MASS)
library(glmnet)

```


***
# CD Algorithm
## Question 3
#### n = 200, p = 10

```{r}
#set parameters for simulation
rho <- 0.9
n <- 200
p <- 10

#create covariance matrix
cov_matrix <- matrix(0, nrow = p, ncol = p)
for (i in 1:p) {
  for (j in 1:p) {
    cov_matrix[i, j] <- rho^abs(i - j)
  }
}

#simulate data
set.seed(100)  # for reproducibility
X <- MASS::mvrnorm(n = n, mu = rep(0, p), Sigma = cov_matrix)

#set b
b <- c(1, rep(0, p - 1))

#lambda value
lambda <- 0.01

# run CD algorithm
cd <- cg_cd(X, b, lambda, tol=1e-4, maxit = 1000)

print(cd)
```

## Question 4
#### Vary $\lambda$ from 0.01 to 0.5  

```{r}
lambda_vals <- seq(.01, .5, by=.01)
cd_result <- vector("list", length = length(lambda_vals))

for (i in 1:length(lambda_vals)){
  cd_result[[i]] <- cg_cd(X, b, lambda_vals[i])
  #print(lambda)
}

df <- data.frame(matrix(unlist(cd_result), ncol=length(lambda_vals), byrow=F))
colnames(df) <- lambda_vals

print(df[,c(1,10,20,30,40,50)])
```

***
# ADMM Algorithm
## Question 3
#### n = 200, p = 10

```{r}
#set parameters for simulation
rho <- 0.9
n <- 200
p <- 10

#create covariance matrix
cov_matrix <- matrix(0, nrow = p, ncol = p)
for (i in 1:p) {
  for (j in 1:p) {
    cov_matrix[i, j] <- rho^abs(i - j)
  }
}

#simulate data
set.seed(100)  # for reproducibility
X <- MASS::mvrnorm(n = n, mu = rep(0, p), Sigma = cov_matrix)

#set b
b <- c(1, rep(0, p - 1))

#lambda value
lambda <- 0.01

#ADMM algorithm
admm <- cg_admm(X, b, lambda)

print(admm)
```

## Question 4
#### Vary $\lambda$ from 0.01 to 0.5  

```{r}
lambda_vals <- seq(.01, .5, by=.01)
admm_result <- vector("list", length = length(lambda_vals))

for (i in 1:length(lambda_vals)){
  admm_result[[i]] <- cg_admm(X, b, lambda_vals[i])
  #print(lambda)
}

df <- data.frame(matrix(unlist(admm_result), ncol=length(lambda_vals), byrow=F))
colnames(df) <- lambda_vals

print(df[,c(1,10,20,30,40,50)])
```

***
# Comparison: CD and ADMM  
#### Computational Speed
```{r}
set.seed(100)
n <- 50
p <- 400
X <- matrix(rnorm(n*p), n, p)
b <- rep(0, 400)
b[301:305] <- c(5:1)*2
y <-  X%*%b + rnorm(n)
ynew <-  X%*%b + rnorm(n)  


#lambda value
lambda <- 0.01

cd_time <- system.time( cd_fit <- cg_cd(X, b, lambda)) 
admm_time <- system.time( admm_fit <- cg_admm(X, b, lambda))

re <- system.time(fit <- glmnet(X, y, lambda = 1, standardize  = F, intercept = F))

t = list(cd=cd_time, admm=admm_time, glmnet=re)
print(t)
```

  
  
#### Accuracy  

```{r, include=FALSE}
re.glmnet <- as.numeric(coef(fit))[-1]
results <- cbind( re.glmnet[301:305], cd_fit[301:305], admm_fit[301:305])

colnames(results) = c("glmnet","cd","admm")

print(results)
```
```{r}
lasso_obj <- function(X, y, beta, lambda) {
    (1/2)*(1/length(y))*norm(y - X%*%beta,"F")^2 + lambda*sum(abs(beta))
}

beta_glmnet <- as.numeric(coef(fit)[-1] )
print(data.frame(cd_diff=max( abs( beta_glmnet - cd_fit )), admm_diff=max( abs( beta_glmnet - admm_fit )))) 
```

***
# Bonus
```{r}
#set parameters for simulation
rho <- 0.9
n <- 200
p <- 10

#create covariance matrix
cov_matrix <- matrix(0, nrow = p, ncol = p)
for (i in 1:p) {
  for (j in 1:p) {
    cov_matrix[i, j] <- rho^abs(i - j)
  }
}

#simulate data
set.seed(100)  # for reproducibility
X <- MASS::mvrnorm(n = n, mu = rep(0, p), Sigma = cov_matrix)

#set b
b <- c(1, rep(0, p - 1))

#lambda value
lambda <- 0.01

#ADMM algorithm
elastic_cd <- cg_cd_alpha(X, b, lambda, alpha=.05)
elastic_admm <- cg_admm_alpha(X, b, lambda, alpha=.05)

print(data.frame(cd=elastic_cd,admm=elastic_admm))
```

