---
title: "PH1930 Final Project"
author: "Group 2: Yu Bin Chen, Rachit Sabharwal, Caroline Schaefer, Pagna Sok"
date: "2023-12-10"
output:
  html_document: default
  pdf_document: default
---

This report contains results of minimizing the equation using the coordinate descent (CD) and alternating direction method of multipliers (ADMM). Functions are run using C++ code and implemented with the R packages Rcpp and RcppArmadillo: 

\begin{align*}
\frac{1}{2}u^TAu-ub^T+\lambda ||u||_1\\
\end{align*}  



#### Here we are loading required packages as well as the CD and ADDM C++ source code
```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS)
library(Rcpp)
library(RcppArmadillo)

sourceCpp("coordinate_descent_algo.cpp")
sourceCpp("admm_cpp_v3.cpp")
```



## Stimutaiton run 1

#### In stimulation run 1, we will run using these parameters: $\lambda = 0.01$, $\rho = 0.9$, $n = 200$, and $p = 100$.

```{r}
#set parameters for simulation
rho <- 0.9
n <- 200
p <- 10
#lambda value
lambda <- 0.01


#create covariance matrix
cov_matrix <- matrix(0, nrow = p, ncol = p)
for (i in 1:p) {
  for (j in 1:p) {
    cov_matrix[i, j] <- rho^abs(i - j)
  }
}

#simulate data
set.seed(100)  # for reproducibility
X <- MASS::mvrnorm(n = n, mu = rep(0, p), Sigma = cov_matrix)

#set b
b <- c(1, rep(0, p - 1))


#ADMM  and CD algorithms and compare their results
cd <- cg_cd(X, b, lambda, tol=1e-4, maxit = 1000)
admm <- cg_admm(X, b, lambda, tol=1e-4, maxit = 1000)
```

***
#### Here is the result for CD:

```{r}
print(cd)
```
***
#### Here is the result for ADMM:
```{r}
print(admm)
```


## Stimuation run 2

#### In stimulation run 2, we will run using the same parameters as in run 1 but we will be varying the $\lambda$ value between 0.01 and 0.5.
```{r}
set.seed(100)

#simulation parameters
n <- 200
p <- 10
rho <- 0.9
lambda_values <- seq(0.01, 0.5, by = 0.01)
max_iter=1000

#generate variance covariance matrix pxp
cov_m<-matrix(nrow=p,ncol=p) 
for (i in 1:p){
  for (j in 1:p){
    cov_m[i, j] <- 0.9^(abs(i-j))
  }
}

#simulate multivariate normal observations
X <- MASS::mvrnorm(n, mu = rep(0, p), Sigma = cov_m)

#create the vector b
b <- c(1, rep(0, p - 1))
```


***
#### Here is the result for CD:

```{r}
#testing different lambda values
for (lambda in lambda_values) {
  result <- cg_cd(X, b, lambda, tol=1e-4, maxit = 1000)
  
  cat("\nResults for lambda =", lambda, ":\n")
  print(data.frame(True_Coefficients = b, Estimated_Coefficients = result))
}
```

***
#### Here is the result for ADMM:
```{r}
#testing different lambda values
for (lambda in lambda_values) {
  result <- cg_admm(X, b, lambda, tol=1e-4, maxit = 1000)
  
  cat("\nResults for lambda =", lambda, ":\n")
  print(data.frame(True_Coefficients = b, Estimated_Coefficients = result))
}
```

